```{r,echo=FALSE,message=FALSE,warning=FALSE}
r3dDefaults = rgl::r3dDefaults
r3dDefaults$FOV = 50
r3dDefaults$zoom = 1

knitr::opts_chunk$set(
  comment =  "#>", 
  collapse = TRUE,
  fig.align = "center")

library(lidR)
rgl::setupKnitr(autoprint = TRUE)
```

# Derived metrics at the tree level {#tba}

## Overview {#tba-overview}

The "tree" level of regularization corresponds to the computation of derived metrics centered on each tree using the points that belong to each tree. Derived metrics calculated at tree level are the basis for an inventory at the individual tree level or the basis for individual species identification.

Similarly to what we have seen in sections \@ref(metrics), \@ref(cba), and \@ref(aba) calculating derived metrics is straightforward and works exactly the same way as in `cloud_metrics()` or `pixel_metrics()`. Derived tree metrics are calculated using the `crown_metrics()` function. The input data for this function is a point cloud that needs to contain information from tree segmentation (e.g. usually the `treeID` attribute). In the majority of cases `crown_metrics()` is run after segmenting tree crowns with `segment_trees()` (section \@ref(itd-its)) but the segmentation could also be performed in another way independently of `lidR`.

In the example below we show the basic use of the `crown_metrics()` function on the files we used in section \@ref(itd-its). This file already stores an ID for each point referring to each tree, so we don't need to perform the segmentation first. We compute two metrics  `z_max` and `z_mean` using the formula `list(z_max = max(Z), z_mean = mean(Z)`. The output is a `sf/sfc_POINT`.

```{r}
LASfile <- system.file("extdata", "MixedConifer.laz", package="lidR") 
las <- readLAS(LASfile, filter = "-drop_z_below 0") # read the file

metrics <- crown_metrics(las, ~list(z_max = max(Z), z_mean = mean(Z))) # calculate tree metrics
head(metrics)
```

The XYZ coordinates of the points correspond to the XYZ coordinates of the highest point within each tree and each point is associated to 3 attributes `treeID` + the two user-defined metrics. In the plot below the output is visualized using color to depict the `z_max` metrics.

```{r spplot-tba-metrics}
plot(metrics["z_max"], pal = hcl.colors, pch = 19) # plot using z_max
```

Like other functions seen in sections \@ref(cba) and \@ref(aba), users can create their own custom functions containing all of the metrics of interest. In the example below we show how to calculate metrics that are based both on point heights and intensities.

```{r}
custom_crown_metrics <- function(z, i) { # user-defined function
  metrics <- list(
     z_max = max(z),   # max height
     z_sd = sd(z),     # vertical variability of points
     i_mean = mean(i), # mean intensity
     i_max  = max(i)   # max intensity
   )
   return(metrics) # output
}

ccm = ~custom_crown_metrics(z = Z, i = Intensity)
```

`crown_metrics()` can also return `sf/sfc_POLYGON` by changing the `geom` argument

```{r spplot-tba-custom-metrics-covex, warning=FALSE}
metrics <- crown_metrics(las, func = ccm, geom = "convex")
plot(metrics["z_max"], pal = hcl.colors)
```
```{r spplot-tba-custom-metrics-concave, warning=FALSE}
metrics <- crown_metrics(las, func = ccm, geom = "concave")
plot(metrics["z_max"], pal = hcl.colors)
```

## Applications {#tba-applications}

### Selection of trees {#tba-applications-tree-selection}

`crown_metrics()` gives the ID of trees and associated metrics. We can use these data to filter the scene and remove trees with a low intensity.

```{r plot-las-selected-trees, rgl = TRUE}
metrics <- crown_metrics(las, ~list(imean = mean(Intensity))) # calculate tree intensity metrics
metrics <- metrics[metrics$imean > 80,] # filter intensity

subset <- filter_poi(las, treeID %in% metrics$treeID)
x <- plot(las, bg = "white", size = 4)
plot(subset, add = x + c(-100, 0), size = 5) # some plotting
```

### Tree based inventory {#tba-applications-tree-inventory}

Assuming we know a relationship between the derived metrics and a value of interest `G` - such as the biomass of a tree - we can map the resource. Lets assume that $G = 0.7 \times z_{max} + 0.1 \times i_{mean}$. In real life a value of interest is more likely to be related to the crown size, but this is a simplified example. First we can compute `G` for each tree:

```{r spplot-tba-G}
metrics <- crown_metrics(las, func = ~custom_crown_metrics(Z, Intensity)) # calculate intensity metrics
metrics$G <- 0.7 * metrics$z_max + 0.1 * metrics$i_mean # set value of interest
plot(metrics["G"], pal = hcl.colors, pch = 19) # some plotting
```

Then using a raster package, we can rasterize the map. Here we get the sum of `G` from each tree within each 15 m pixel to get the total value of `G` for a given pixel. The final output is a predictive model that mixes the area-based approach and the tree-based-approach.

```{r plot-raster-sumG, fig.height=5, fig.width=6}

r <- terra::rast(ext(las),  resolution = 15)
v <- terra::vect(metrics["G"])
map <- terra::rasterize(v, r, field = "G", fun = sum) # extract sum of G at 15m
plot(map, col = hcl.colors(15)) # some plotting
```

This is a small example that may not be of interest, but one may imagine or test the result on a bigger data set.
